\textbf{Unshuffling Permutations}

### 2016-04-06 slides/slides.tex
melanges

\documentclass[unknownkeysallowed,11pt]{beamer}

@@

\DeclareMathOperator{\QQ}{\mathbb{Q}}

@@

>>>>

\documentclass[unknownkeysallowed,10pt,xcolor={dvipsnames}]{beamer}

@@

\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{std}}
\DeclareMathOperator{\SHUFFLE}{\bullet}

\newcommand{\BIB}[2]{{\footnotesize\textcolor{MidnightBlue!85}{[\textbf{#1}, #2]}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Example: unshuffling binary words}
Let $B$ the set of all binary words and let us consider the coproduct
\begin{equation*}
    \Delta : \QQ[B] \to \QQ[B] \otimes \QQ[B]
\end{equation*}
on binary words defined by
\begin{equation*}
    \Delta(u) :=
    \sum_{P \sqcup Q = \{1, \dots, |u|\}}
    u_{|P} \otimes u_{|Q}.
\end{equation*}

\begin{block}{Example}
\begin{equation*}
    \Delta(100) =
    \epsilon \otimes 100 + 1 \otimes 00
    + 2 (0 \otimes 10) + 2 (10 \otimes 0) + 00 \otimes 1
    + 100 \otimes \epsilon
\end{equation*}
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Example: shuffling binary words}
\begin{block}{Proposition}
    The shuffle product of binary words is the dual product of the
    unshuffling coproduct of binary words.

        REF ? SCHUTZ ?
\end{block}
\medskip

\begin{block}{Example}
\begin{equation*}
    \Delta(100) =
    \epsilon \otimes 100 + 1 \otimes 00
    + \textcolor{BrickRed}{2 (0 \otimes 10)} + 2 (10 \otimes 0) + 00 \otimes 1
    + 100 \otimes \epsilon
\end{equation*}
\smallskip

Since $0 \otimes 10$ has multiplicity $2$ in $\Delta(100)$, by duality,
$100$ has multiplicity $2$ in $0 \shuffle 10$:
\begin{equation*}
    0 \shuffle 10 = 010 + \textcolor{BrickRed}{2 (100)}.
\end{equation*}
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Unshuffling permutations}
If $u$ is a word of integers without multiple occurrence of a same letter,
the \alert{standardized} $\STD(u)$ of $u$ is the unique permutation of
$S_{|u|}$ order-isomorphic to~$u$.

\begin{block}{Example}
\begin{equation*}
    \STD(82194) = 42153
\end{equation*}
\end{block}
\medskip

\begin{block}{Definition}
    The \alert{unshuffling coproduct of permutations} is the
    coproduct on $\QQ[S]$ defined by
    \begin{equation*}
        \Delta(\pi) =
        \sum_{P \sqcup Q = \{1, \dots, |\pi|\}}
        \STD\left(u_{|P}\right) \otimes \STD\left(u_{|Q}\right).
    \end{equation*}
\end{block}

\begin{block}{Example}
\begin{equation*}
    \Delta(213) =
    \epsilon \otimes 213 + 2 (1 \otimes 12) +
    1 \otimes 21 + 2 (12 \otimes 1) + 21 \otimes 1
    + 213 \otimes \epsilon
\end{equation*}
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Shuffling permutations}
\begin{block}{Definition}
    The \alert{shuffling product of permutations} is the product
    $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the unshuffling
    product of permutations.
\end{block}

\begin{block}{Example}
\begin{equation*}\begin{split}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
    2134 + 2 (2314) \\
    & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
    3 (3214) + 2 (3241) \\
    & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
\end{split}\end{equation*}
\end{block}
\medskip

This shuffle product has been introduced for the study of some
Hopf algebra \BIB{Vargas}{2014}.
\medskip

It is a kind of association of the convolution product and the shifted
shuffle product of permutations \BIB{Duchamp, Hivert, Thibon}{2002}.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}
